
\begin{description}


\item[Schedulable:] A couple of hours on weekday evenings is enough. Play online with vtt and voice. Simple, quick, no fuss. 
\textbf{\textit{Just Get Gaming!}} 

The game is explicitly designed to play online, so busy people can play. We have had hundreds of sessions since 2008, even when people have been spread out over different countries, had demanding careers, lots of travel, young kids, etc.


\item[Kids:] Starts simple enough for 5yo kids, and scales up through university. Sparkly Pony, Lightning McQueen, and the old Rubber Troll make excellent Heroes as they go explore the dark depths below the dining room table.

Or play with pen and paper, meeples or minis, battle maps or improvised terrain. You don't need much to start, and you already have it at home.


\item[Hundreds of hours of adventures:] Built and tested over many years. Loads of adventures and several campaign arcs present highly varied experiences, stories, and challenges, all across Kingsland. 


\item[Share and meet new gamers:] Upload and share your adventures with the greater community. Find new friends, steal cool ideas, adventures, characters, maps, figs, etc, discuss that \emph{awful} rule I just pushed, and how much better it would be \emph{this other way}.

\

\item[Fast:] Lots of quick rounds. Attack, retreat, sneak around behind, flank, make a stand until the sorcerer has finished his summoning, then disengage and escape through the secret tunnel. All in a fluid and varied battle in the dungeons of the Antorg Slavers.

Supports 50+ Heroes and Monsters on the map in full glorious slaughter, but keep it below 20 and no one will get bored.


\end{description}
