%--------|---------|---------|---------|---------|---------|---------|---------|
%       10        20        30        40        50        60        70        80




%-------------------------------------------------------------------------------
%M A G I C   S K I L L S
%-----------------------

\cleardoublepage

\phantomsection\addcontentsline{toc}{chapter}{Magic}
\chapter*{Magic}
\chaptermark{magic}
\label{cpt:magic}


\phantomsection\addcontentsline{toc}{section}{skills}
\section*{Magic skills}

All magic skills and spells must be learned from a teacher, book, or similar. They cannot be trained from nothing without guidance like normal skills. This includes the base skills \emph{magic} and \emph{spell caster}, as well as the support skills like \emph{power casting}, and all spells.

Teachers can be the village witch, a professor at the City Magic College, an ally in the party who happen to know that spell, or the enemy wizard you captured a few weeks ago and been torturing ever since (\emph{but would you trust him?}). When creating a new character with magic skills and spells it's assumed that he had access to some magic training when growing up, perhaps 1d3 spells.

Some magic items can be used without any magic skills, others require that the user has some knowledge of the \emph{magic} skill. To learn and cast spells it is necessary to have trained \emph{spell caster}, and that is least expensive directly when the character is created.


\begin{description}


\item[Skill scf 2.0: "magic":] The caster can not spend more mana on each spell casting than his level in magic. This also limits the amount of mana he can pump into an already active spell each round, or how much mana he can use to activate or power a magical item.

The magic skill is also used when trying to understand magical texts and foreign spells and devices, etc. Or when activating certain magical items. It also provides some protection against certain types of spells.

Note: Not all magical devices, gadgets, trinkets, and weapons require the magic skill to use. Some just require being drawn, calling an activation word, or other circumstance. Some gadgets will also draw mana from the user by themselves without the user needing the magic skill to infuse them with power, and some items are powered by themselves or other sources.


\item[Skill cost special "spell caster":] is requried to be able to learn and cast spells. It is not required for the use of magical items nor to train the magic skill.

The xp cost to train spell caster varies depending on the total xp of the character, increasing as the character advances:\\
If trained at character creation it costs 20xp.\\
After that it costs 20 + (total xp / 10).\\
E.g: For a 123xp Hero it costs 33xp, and for a 358xp Hero it costs 56xp to train spell caster.


\item[Skill scf 1.0: "power casting":] is when the spellcaster pays extra mana power to increase his chances for the casting roll. The caster gains mod+1 for each mana spent. The caster can not spend more mana than his skill in power casting. This cannot be used to raise the end total chance of success above 9.

Mana spent for power casting is not counted to the maximum total limit set by the caster's magic skill level.


\item[Maneuver cost 10: "held spell":] delays the final touch of the spell so that the effect can appear on very short notice. Can with good effect be used together with anticipate to inject the effects of a prepared multi-round-casting spell at a precise point in the action sequence of a battle round. It costs 3ap to cast a held spell. Holding a spell is free in the first few rounds but costs 1 mana per every exceeded period of 1 + int/3 rounds.

\

\todo fix "pinpoint on map" issues, miss allowance distance = lvl squares.

\item[Maneuver cost 5 + scf 0.5: "indirected strike":] allows the casting of \emph{strike, chock, flash} spells without direct line of sight to the target. By taking a mod-3 or 1r extra casting time the wizard can set a target for the spell by an approximate location, such as a description from a spotter communicated in 1 action (3ap), guesstimated location from recently seeing the target move out of sight, and similar, GM discretion.


\end{description}






%-------------------------------------------------------------------------------
%S P E L L S
%-----------

\phantomsection\addcontentsline{toc}{section}{spells}
\section*{Magic spells}

Magic spells cannot be trained from nothing without the user learning the spell from a master, scroll, grimoire, etc. Purchasing a spell text costs 1-10 gold depending on the rarity, strangeness, power of the spell, etc. GM discretion.
Spells can also be deduced from magical objects and other sources, with a lot of study and some modified magic rolls.

Consider the spells listed below to be suggestions only. Build individual spells to make the use of magic more varied and unreliable. Perhaps the Hero wants to purchase a Black Bolt spell, but the grey old master only teaches him the Thread of Dread Darkness which is very similar to the Black Bolt but with some minor changes in range, effect, etc.

Magic spells are generally slow to cast. A normal time spell takes 1-2r to cast, a slow spell can take 3r or longer. A fast spell takes a couple of actions or one round.

When casting a spell, each mana pumped into the spell, including casting cost, reduces the chance of success by one. I.e. a black bolt with two extra mana for damage and one for range will cost 4 mana to cast and carry a mod-4 to cast.

Spells that have range don't allow for short,long,vlong,extreme range modifications of any sort unless explicitly stated. E.g: a black bolt r10 will not reach a target 11sq away from caster, but a shock bolt r20 can reach 30sq at dam-3.


\begin{description}

\item[Skill scf x.x: \emph{spell}] Most spells have a default skill cost factor of 0.33. Some spells are cheaper, some more expensive. Some spells have minimum requirements on psy or int. 

If the character does not meet the int requirements, then the cost of the spell is doubled for each missing int point: e.g: Wizard Woolhead has int 4 and wants to learn an int 7 spell. He is missing 3 int points, thus the spell costs $2^3=8$ times as much to learn. He then spends his xp on a simpler spell.

If the character does not meet the psy requirements he will have mod-1 per missing psy when casting the spell.

\end{description}

These spells should be seen as examples, different masters teach their students different variations, some better, some worse, some just different. It is likely that most wizards' spells will be different from each other.

Spells with more complexity or targets have higher int requirements, and spells with higher power or control influence has higher psy requirements.

\subsection*{Custom spells, player created spells}
%-------------------------------------------------
More fun is to let the player write new spells for his character, and you as GM put requirements on them. That way the spell casters become more unique, and it's more personal for the player. This could also be a way around the limitation that the spells have to be tought by someone. Perhaps the Hero can just figure it out by himself?

When the spells are new they are in flux. Prototypes. And they have to be tested under a wide range of conditions. The first few adventures and tens of usages of the spells in varied situations should give you as GM, the player, and the rest of the group a decent insight into the functions and aspects of the spell. During this period the details, stats, and requirements of the spell should be flud and change over time, until you've found the sweet spot. Remember to consider how the spell might work if it's used by a different caster, or in combination with a different group, tactics, opposition, equipment, combat styles, etc.






% typical: relative base, 1h/2h  dam , extras
% sword    +2   +2/+4                             finesse 6
% axe      +3   +3/+6                             parry-3 toparry-1 toavoid+1
% spear    +1   +1/+2 pen 0/1 parry -1/+1         reach
% staff    -1   -1/+1 parry +1/+2                 fast, reach
% braw     -3                                     (at str 6)
%   fist   -4                                     fast, deflect
%   kick   -2                                     deflect
%
% bow      =0   --/=0                             20sq, pen, 1/r
% crossbow +3   --/++ pen 1-3                     15sq, pen, 1/2r
%
%
% typical psy int =6, mana 15
%             assume armour2 for ppp
% blkblt1  =0 +2p                                 10sq, ppp, 1/r
% blkblt2  +3 +5p                                 10sq, ppp, 1/2r
% shkblt1  =0 +2p                                 15sq, ppp, 1/r
% shkblt2  -1 +1p                                 20sq, ppp, 1/2r
% frcblt3  -1 =0p                                 10sq, p1,  3ap
% frcblt3  -3 -1p                                 10sq, p2,  3ap


% guide lines ?
% psy - power, damage, effect
% int - speed, complexity, range, multi-target, control
%
% assume baseline psy+3 dam+pen, count full penetration as pen2 on average
% ok to have psy+3 dam+pen + scale (psy+3)/4 per mana for 1-2r casting
% range 10 as baseline, trade 1dam/pen for 5 range
% usually no extended ranges, all based on one max range, some with long etc
% 








% baseline for spells:   int 3, psy 3
%                        cast 1r 1m          spear/strike, +1r casting for bolts
%                        dam 6 (5-7) pen 0
%                        range 10
%
%  int +1 for range extension flexibility
%  int +1 for damage extenstion flexibility
%  int +3 for range+5          or trade dam + cast time, see below
%  int +3 for quicker casting  3r > 2r,   2r > 1r,   1r > 3ap
%
%  psy +1 for dam+1 / pen+1 / more power / effect
%  trade 3 dam+pen against 5 range






% damage psy based, parity with sword, alpha boost (dam/3) / mana
% medium range as base range for bolts
% no parry or avoid defense against bolts
% no shield defense or cover for bolts
% no shooting into melee mods
% no movement speed mods
% 
% base dam  psy+2
% base range 10
% base cast 2r for spear and strike, 3r for bolt
% 
% dam class time    base int   dam/psy
% dam 1-3    1a/3ap  int6       dam = psy    4ap : dam=psy+1 or extra effect ?
% dam 4-6    1r      int5       dam = psy+1  or psy=dam +pen1
% dam 7-9    2r      int4       dam = psy+2  or psy=dam +pen2
% dam 10-12  3r      int3       dam = psy+3  or psy=dam +pen99
%
% ? but can't generally scale base damage higher than (psy dam) / mana
% ? even at long times
% 
% treat penetrating as dam+3 on average,   trade dam <> pen equally
%
%
%
%
% scale extra mana for dam / range / stun / radius / pen / etc
%
% damage:                    int base    dam base
% 3ap  dam 1-3    +1/m         int6      dam psy
% 1r   dam 4-6    +2/m         int5      dam psy+1
% 2r   dam 7-9    +3/m         int4      dam psy+2
% 3r   dam 10-12  +4/m         int3      dam psy+3
%
% range:                     int mod AND dam mod
%      touch                   int-3     dam+3
%      r4sq       +2/m         int-2     dam+2
%      r7sq       +3/m         int-1     dam+1
%      r10sq      +5/m         int       dam
%      r15sq      +7/m         int+1     dam-2
%      r20sq      +10/m        int+2     dam-3
%      r25sq      +12/m        int+3     dam-4

%
% treat bolt as one class higher than spear or strike :  more useful, no target mods
% treat ball as two classes higher than spear and strike  :  are effect, no mods
% 
%
% Trade cast time and range against damage and psy int requirements
% DONT trade more than one time class !!!  NOT 3r > 2r > 1r for int+6 psy+6
%
% cast 3r > 2r        dam-3 /  psy+3 & int+3
% cast 2r > 1r        dam-3 /  psy+3 & int+3
% cast 1r > 3ap       dam-3 /  psy+3 & int+3
% 
% range 10 > 15       dam-2 / cast+1r
% range 10 > 20       dam-3 / cast+1r
% allow long @dam-2   dam-1
% area effect         dam-2 per radius+1  /alt  cast+1r per radius
% 
% 
% int 3 as base for normal complexity
% int+2 for multi target
% int+2 for area effect, int+1 per extra radius+1
% int+1 for extra range / mana
% int+1 for extra damage / mana
% int+1 for pen1-2, int+2 for pen3-ppp
% int+1 for buffer / hold effect
% trade int-1 for psy+1 or mana+1 cost
% trade int-1 for +1r cast time
% trade range-5 for int-1
% trade range touch for int-3 ?
%
% 
% Examples below are BOLTS and thus +1r casting
% comparable spear/strike should be   3r >> 2r,   2r >> 1r,   1r >> 3ap
%
% black bolts  cast 3r  !! dam 9            >> int=3, psy=7
%              dam psy+0 dam+/m ppp         base psy+2, dam-2 ppp, int+1
%              range 10 (med)               int,psy +0
%              range +5/m                   int+1
%              >>                           int 6, psy 9
%
% shock bolt   cast 3r  !! dam 9 fundament  range 20  dam-3 
%              dam psy+0 dam+/m ppp         base psy+2, dam-2 ppp, int+1
%              range 20                     range 20  dam-3
%              range long @dam-3            dam-1, but take int+2 instead ?
%              range+/m                     int+1
%              >>                           int 8, psy 9
% 
% force bolt   cast 1r      (2r>>3ap : dam-3, psy+3 & int+3)  psy+2,-3=psy-1
%              range 10                     --
%              
%
%
%
%
% stun effects ?   
%
% add stun:   int+1,  stun3 = dam-1





%-------------------------------------------------------------------------------
% test spellslist formatting

%\
%
%\
%
%\todo testing spellslist formatting
%%-------------------------------------------------------------------------------
%%\TODO remove or switch to this:   check how it looks with spells as list ------
%\openspellslist
%
%\spell{Ur Bolt} int 3, psy 3, scf 0.33, \\
%cast 2r 1m, 
%dam 6, 
%range 10 \\
%\textit{Of Ebedan Ur, oldest of bolts, archetypical of the ancient spellcasting arts of killing. A quick flash of clear blue light, sharp crack, smell of dry parchment. But when cast less eloquently it's well known to produce the soft quack of a drunk duck and a faint smell of old wine.}
%
%\spell{Onsager Spear} int 6, psy 7, scf 0.33, \\
%cast 3r 1m, 
%dam 10 +3/m, 
%range 10 +5/m \\
%\textit{The Onsager battle monks developed a simple martial spellcasting creed, and the Spear is their most classic, simple, and well known infantry battle spell. It's unknown how they managed to bring their massed infantry to the required mental capacity.}
%
%\
%
%\spellslistnote{Jeremiah Black, with no imagination and well known to never drink anything stronger than the milk of his two headed goats. Survived only by his most traditional, efficient, and horribly boring Black Bolts. They tend to have high requirements but have been shown to be highly effective in a wide range of situations.}
%
%\spell{Black Bolt} cast 2r 1m, 
%dam 9 +3/m penetrating,
%range 10 +5/m,
%int 6 psy 7 \\
%\textit{A faster cast, medium power version, of Black's bolt library.}
%
%\spell{Black Bolt} int 6, psy 9, scf 0.33, \\
%    cast 1r 1m, 
%    dam 6 +2/m penetrating,
%    range 10 +5/m,
%    int 6 psy 9\\
%    \textit{A weak but very fast cast version of the archetypical Black Bolt.}
%
%
%\closespellslist
%%-------------------------------------------------------------------------------
%
%\
%
%\






\subsection*{Bolts, lances, missiles, strikes, sparks, etc}
%----------------------------------------------------------
Many offensive spells have similar general mechanics. 

\begin{description}

\item[bolt, zap, spark, lightning:]
Bolts don't take target \emph{to hit} modifications on the spell casting roll. They ignore things like target partial cover, shields partially in the way, target movement speed, partially obstructed view by fire, etc.
Generally these magic energies don't travel in a straight line from caster to target but might curve, form a branched lightning tree, split and reform, or snake forward erratically.

These spells are more powerful than the spear or strike categories and often require more time or mana to cast for similar effect.


\item[spear, lance, dart, arrow, quarrel, missile:]
These work similar to projectiles. The spells take target \emph{to hit} mods directly as casting mods. Target size, target movement speed, partial cover, shields in the way, shooting into melee, fire, etc all mod the spell casting roll directly. These are the same to-hit target mods as a firing arrows from a bow would take. 

%TODO: think there are edge cases where this will not work right

Often the spell formation shoots in a rapid line to hit the target, or whatever gets in the way. A failed casting, which would have succeeded if not for the \emph{to hit} mods from cover, movement, target size, etc can be interpreted as a successfully cast spell which strikes cover or near the target instead.

%TODO: should this then still draw full mana, or reduced mana for minor fail? Can draw reduced mana and still make a nice flashy effect.

Since the spell behaves like a projectile, taking mods like a bow or crossbow, skills like sniper, lead target, etc work to alleviate those mods, just like for bows.


\item[strike, chock, flash:] 
These are mental target lock spells that take casting mods based on target psy, actions, and magic training. Thus they tend to be less useful against other spellcasters. The casting roll takes the following mods, stacking, based on target:\\
mod -(target psy / 3)\\
mod -3 if target takes a 3ap action to resist the spell\\
mod -(target magic lvl / 3) (rd) E.g: magic 4 : mod-2
% mental lock spells ignore cover etc, but mods by psy etc
%              ignores target move mods, shooting into melee mods
%              BUT: they take cast mods for target psy, if target resists,
%                   and if the target knows magic skill.
%                   mod -(target psy / 3), 
%                   mod -3 if target takes a 3ap action to resist 
%                   mod -(target magic lvl) 
%              Hence, almost useless against wizards


\item[blast, ball, cloud, storm:]
These are area effect spells. They can be targeted on characters, objects, or locations, and take no target mods.


\end{description}

\

\small \begin{verbatim}
Ur Bolt        cast 3r, 1m, dam 5,
               range 10
               int 3, psy 3

Onsager Spear  cast 2r, 1m, dam 10 +3/m,
               range 10 +5/m
               int 5, psy 8

Keregt Strike  cast 2r, 1m, dam 8,
               range 10 +5/m
               int 4, psy 6

---------------

black bolt     cast 3r 1m, dam 9 +3/m, penetrating,
               range 10, 
               int 5, psy 9

black bolt     cast 2r 1m, dam 6 +2/m, penetrating,
               range 10,
               int 5, psy 9

shock bolt     cast 3r 1m, dam 5 +2/m, penetrating,
               range 20 +10/m, long dam-3,
               int 8, psy 8

shock bolt     cast 2r 1m, dam 4, penetrating,
               range 15 +7/m, long dam-2,
               int 6, psy 6

force bolt     cast 1r 1m, dam 3
               range 10
               int 7, psy 4

force bolt     cast 1r 1m, dam 5 +1/m
               range 10 +5/m, long -2dam,
               int 10, psy 6

---------------

bright arrow   cast 2r, 1m, dam 6
               range 10 +5/m
               int 4, psy 4

bright arrow   cast 1r, 1m, dam 5
               range 10 +5/m
               int 7, psy 7

snuppel dart   cast 2r, 1m, dam 4, stun 5 +3/m
               range 15 +7/m
               int 6, psy 5


---===### current above ###===---
\end{verbatim} \normalsize

\

\noindent 
The spells listed below are not re-balanced yet.\\
See balance guidelines in comment segment above.

\

\small \begin{verbatim}
---===### legacy  below ###===---


fire ball      cast 3r 2m, dam 5 +1/m, range 10 +5/m, radius 2 +1/m,
               duration 3r +2/m, damage each round
               int 8, psy 8

fire ball      cast 2r 2m, dam 4 +1/m, range 10 +5/m, radius 1 +1/m,
               duration 1r +1/m, damage each round
               int 7, psy 6

fire blast     cast 1r 2m, dam 3 +1/m, range 5 +5/m, radius 1 +1/m,
               int 6, psy 6

fire storm     cast 1r 1m, dam 5 +2/m, range self, radius 1 +1/m,
               duration 3r +3/m, damage each round,
               caster is immune to fire for the duration
               The storm follows the caster at most up to walking speed,
               if the caster moves faster the storm will disappear.
               The max damage of the storm each round is reduced by one for
               each movement speed declared for the round.
               int 6, psy 8

minion missile cast 1r 1m, dam 4 +2/m, speed 20, seeking 10 -2/r,
               penetrating 1, lifetime 20r +10/m,
               Hover by caster until target assigned,
               then seeks target until missed seeking or max lifetime.
               Seeking does not fail on roll of 10 as normal.
               They are smart enough to curve around objects and avoid obstacles
               unless they miss their seek roll (one per rnd). The target does
               not have to be in line of sight from the caster once assigned.
               Seeking penalty only after target assignment, not total time
               since summon.
               Assigning one or more bolts to one target is a
               very fast action (1ap).
               int 6, psy 7

ward shield    cast 2r 2m, range touch, duration 5r +3r/m
               charge 5 +3/m (pay on cast)
               Absorbs incoming damage before it strikes the character
               or his equipment. Each absorbed damage point costs
               one charge from the spell.
               Ignores penetrating, i.e. penX just passes through the spell.

ward shield    cast 1r 2m, range touch, duration 5r +3r/m
               charge 4 +2/m (pay on cast)
               Absorbs incoming damage before it strikes the character
               or his equipment. Each absorbed damage point costs
               one charge from the spell.
               Ignores penetrating, i.e. penX just passes through the spell.

ward flash     cast 3ap 1m, range personal, duration 1r first attack only,
               absorbs 5dam +3/m on the first attack that hits the caster.
               Absorbs incoming damage before it strikes the character
               or his equipment.
               Ignores penetrating, i.e. penX just passes through the spell.
               int 5, psy 5

ward skin      cast 2r 2m, range touch, duration 5r +3r/m
               charge 4 +2/m (pay on cast)
               Absorbs incoming damage before it does hp damage, but does not
               spend charges on damage that is absorbed by equipment or armour.
               Each absorbed damage point costs one charge from the spell.
               Ignores penetrating, i.e. penX just passes through the spell.
               int 7, psy 5

force shield   cast 3ap 1m, range personal, duration 1r,
               absorbs 3dam +1dam/m on every attack during the round.
               Absorbs incoming damage before it strikes the character
               or his equipment.
               Ignores penetrating, i.e. penX just passes through the spell.
               This ward absorbs damage for the active round. It is not limited
               in how much damage it can absorb.
               int 5, psy 7

force sphere   cast 2r 1m, range personal, duration 5r +3/m,
               radius self +1/2m.
               Creates a force wall sphere around the caster and anyone inside
               the encompassing radius.The sphere will absorb damage from all
               incoming physical attacks up to a maximum of caster psy per
               attack. The damage absorption costs mana, which is paid per
               attack, and not on casting. Each int/3 damage costs 1 mana.
               Attacks doing more damage than the caster's psy will penetrate
               the shield and strike the target with reduced damage.
               Ignores penetrating, i.e. penX just passes through the spell.
               The caster can cancel the shield at any time ignoring initiative.
               Cancelling the shield is a 0a instant interrupt action.
               Anyone inside the shield can cast spells and attack as usual.
               All actions are mod-1 when inside the shield. Movement costs
               double.
               int 5, psy 7

strike blast   cast 2r 1m, range touch, duration 5r +3/m
               damage 4 +3/m (pay on cast) add to max weapon damage.
               Enchant a weapon for extra damage for the next strike it hits.
               Only active for one strike. Also works on projectiles.
               The weapon takes equal magical damage from the blast,
               check against the abs of the weapon as usual for incoming damage.
               Add +1m to allow a trigger word instead of first strike.
               int 3, psy 5

strike blast   cast 3ap 1m, range touch, duration 1r
               damage 4 +3/m (pay on cast) add to max weapon damage.
               Enchant a weapon for extra damage for the next strike it hits.
               Only active for one strike. Also works on projectiles.
               The weapon takes equal magical damage from the blast, check
               against the abs of the weapon as usual for incoming damage.
               int 7, psy 5

encharge       cast 2r 2m, range touch, duration 5r +3/m
               damage +1 +1/m (pay on cast) add to max weapon damage.
               to hit +1 +1/m (pay on cast) add to chance to hit when attacking.

encharge       cast 2r 2m, range touch, duration 5r +3/m
               abs +3 +3/m (pay on cast) add to weapon abs.
               parry +1 +1/m (pay on cast) add to chance to parry.
               Only works on weapons and shields, not armour.

slow           cast 1r 1m, +1target/2m, range 15 +5/m, duration 5r +2/m
               Movement costs double movement points.
               All actions are slow-2 (takes +2 extra ap)
               Full/multi-round actions take +1r
               Continue slow on psy+2/m vs psy each round (mana paid once
               on cast) up to the max duration. First round auto success.
               Once slow is lost on a target that target is free from
               the spell.
               int 4, psy 3

hasten         cast 2r 2m, +1target/3m, range touch, duration 3r +1/m
               Adds 3ap +1/m to target each round for the duration.
               Adds movement: M+1mp, W+2mp, R+3mp, D+4mp
               Initiative is raised by 3+3/m.
               int 8, psy 8

hold           cast 1r 2m, +1target/2m, range 10 +5/m, duration 3r +1/m
               Target cannot do movement at all, but stationary actions can
               can be performed without penalty.
               Continue hold on psy vs psy +2/m each round (mana paid
               once on cast)
               int 6, psy 7

paralyse       cast 2r 2m, +1target/2m, range 10 +5/m, duration 3r +2/m
               Target cannot move or take actions that require any movement,
               even if it is only with the hands.
               Continue paralyse on psy-3 vs psy +2/m each round
               (mana paid once on cast).
               int 7, psy 9

paralyse       cast 2r 2m, psy vs psy +1/m, range 5 +2/m,
               Target cannot move or take actions that require any movement,
               even if it is only with the hands.
               duration diff+1r +2r/m (pay on cast, diff from psy vs psy roll)
               int 7, psy 8

strengthen     cast 2r 2m, +1target/2m, range touch, duration 5 +2/m
               The target gets strength+3 +2/m for the duration of the spell.
               int 5, psy 5

fortify        cast 2r 2m, +1target/2m, range touch, duration 5r+2/m
               The target gets constitution+3 +2/m for the duration of the spell
               int 5, psy 5

fear           cast 1r 1m, psy vs psy +2/m, range 5 +2/m, area aura,
               duration 2*diff +3r/m
               int 4, psy 6

blind          cast 1r 1m, range 15 +5/m, duration 5r +2/m
               target gets vision=1 for the duration.
               int 4, psy 6

open door      cast 3ap 1m, range 15 +10/m, targets +1/m
               strength 5 +3/m
               Open a door. If the door is locked or barred the strength of the
               spell works as either physical strength to break down the door
               or as "pick lock" to try to open the lock.
               scf 0.2

close door     cast 3ap 1m, range 15 +10/m, targets +1/m
               Close a door.
               If caster has seen the door before he doesn't need line of 
               sight if he can pass an int-3 roll.
               scf 0.2

seal door      cast 3ap 1m, range 15 +10/m, targets +1/m, duration 10r +5/m
               strength 5 +3/m
               Seal a door with the strength of the spell. Attempts to break
               down the door will have to overcome the strength with str vs str.
               scf 0.2

dislocate      cast 2r 1m, range 20 +5/m, target 1 +2/m, psy vs psy +3/m
               target forgets the local location information.
               (Press ctrl+shift+o in maptools to reset map info to what
               the characters can see at the moment, as long as all heroes
               have been successfully dislocated).
               int 5, psy 5

pilfer         cast 2r 1m, range 5 +5/m,
               Can steal one item from the target.
               If the item is unseen (e.g: in a container) than it is a random
               item from among the potential items.
               A held item can be stolen with a successful psy-6 vs psy +1/m
               roll.
               The item vanishes from the old location and materialises in the
               caster's hand, with a nice tingling pop sound.
               int 5, psy 6

heroism        cast 2r 1m, range 10 +5/m, duration 10r +5/m
               select up to 5 +3/m targets which get that heroic feeling
               they receive +6 psy to all fear tests
               they receive +1 to all offensive actions
               The effect vanishes if the targets get outside the range of the
               spell, which is centred on the caster for the duration.
               A target looses the effect if he does not attack in a round
               when he has the opportunity to do so, unless attacking is clearly
               detrimental to his goal.
               int 5, psy 5

heal           cast 2r 1m, heal 3 +2/m, time 2hp/r, range contact
               int 5, psy 4

heal           cast 2r 1m, heal 2 +2/m, time 3hp/r, range contact
               int 7, psy 6

heal           cast 3r 1m, heal 5 +3/m, time 1hp/r, range contact
               int 7, psy 7

heal           cast 1r 1m, heal 1 +1/m, time 3hp/r, range contact
               int 6, psy 6

heal           cast 2r 1m, heal 1 +1/m, time 5hp/r, range contact
               int 9, psy 7

heal all       cast 3r 1m, heal 1 +1/m, time 2r/hp, range 3 +1/m
               heals all creatures within range.
               int 7, psy 7

force wall     cast 1r 1m, size 3 +1/m, duration 10r +5/m
               psy vs str +3/m to stop anyone passing the blocked passage
               attacks and thrown or shot objects are always blocked
               int 3, psy 6

force wall     cast 2r 2m, size 3 +1/m, duration 10r +5/m
               psy vs str +3/m to stop enemies passing through the blocked
               passage, friendlies can pass at maneuver speed
               attacks and thrown or shot objects are always blocked
               int 8, psy 7

staff light    cast 2r 1m, duration 50r +25r/m
               can be other objects than staff, but a trusty object is mod+3.
               basic light intensity is candle (5)
               +1m gives light intensity as lamp (10)
               +2m gives light intensity torch (15)
               +3m gives light intensity fire (20)
               +4m gives light intensity strong (30)
               int 3, psy 2, scf 0.2

marsh light    cast 2r 1m, duration 20 +20r/m
               Create a candle intensity orb of light that the caster can
               move around at speed 3 +3/m. Moving the light any amount in one
               round is a fast action pushing mod-2 to ams.
               +1m gives light intensity torch (10)
               +2m gives light intensity fire (20)
               +3m gives light intensity strong (30)
               int 5, psy 3, scf 0.2

mana transfer  cast 2r 1m, transfer 3m/r +3/m, range 5,
               willing target or object only, in/out
               Transfer is full round action, can only maneuver
               and do no actions for the duration.
               Max psy mana per cast.
               int 3, psy 3

raise dead     cast 5r 2m, duration 10r + 3*diff +10r/m, range touch
               Immediately activates control dead, with same diff, if the
               caster knows the control dead spell. If not ...
               int 4, psy 4

control dead   cast 2r 1m, psy vs psy +1/m +diff
               duration 3*diff +10r/m, range 20 +10/m
               Activates automatically after successfully raising a dead.
               Must roll psy vs psy +1/m +diff each round to keep control,
               mana spent once when casting or activating.
               When control is lost it can be regained by casting again.
               The caster is aware of his minion's surroundings as though
               being there. Vision is min(psy_caster|psy_target).
               Giving a control directive takes 1a, and must be very simple,
               e.g: attack target/group, move to location, guard area, etc.
               int 8, psy 8

lower dead     cast 3r 1m, psy vs psy +1/m, range 5 +5/m
               int 5, psy 7

mass raise     cast 10r 3m, duration 10r + 3*diff +20/m, range 5 +5/m
               target diff +1/m
               int 8, psy 8, scf 0.66

mass control   cast 3r 2m and psy-3 vs psy +1/2m +diff
               duration 3*diff +10r/m, range 30 +10/m
               Activates automatically after successfully raising the dead.
               Must roll psy vs psy +1/2m +diff each round to keep control,
               mana spent once when casting or activating.
               Simplify if necessary to: duration is reduced by the chance of
               breaking free squared.
               When control is lost it can be regained by casting again.
               The caster is aware of his minion's surroundings as though
               being there. Vision is min(psy-caster|psy-target).
               duration 3*diff +10r/m, range 50 +20/m
               int 10, psy 10, scf 0.66

mass lower     cast 5r 2m, psy-3 vs psy +1/2m, range 5 +5/m
               int 8, psy 8, scf 0.66

drain          cast 3r 1m, psy vs psy +3/m, range contact,
               drains diff mana points from target
               gives caster diff/2 (round up) mana, i.e. half is lost
               Must be in contact during whole casting.
               Cannot drain the target below 0 mana. If draining the target to
               0 mana the target must roll psy or falls unconscious. Roll psy 
               once per round to wake up again.
               int 5, psy 8

darkvision     cast 2r 1m, vision 15+5/m arc 180+90/m, duration 20r+20r/m
               int 3, psy 2

rot            cast 2r 1m, dam 4 +1/m 1hp/r penetrating, range 8 +2/m,
               magical rotting disease is doing the damage.
               one pain per damage dealt.
               int 6, psy 5

poison gas     cast 3r 2m, , area radius 4 +1/m, duration 10r +5r/m
               dam 1/r penetrating. Damage only if psy vs con +diff +3/m,
               roll once for susceptibility.
               int 4, psy 4

revive         cast 5r 5m, psy vs psy +1/m, the dead wanna stay dead.
               the body must first be healed to at least 1 hp before or during
               the revival ritual.
               The character tends to come back altered from the ordeal in the
               world of the dead.
               int 8, psy 10, scf 1.0

fumbly         cast 2r 1m, duration 5r +2/m, range 5 +2/m, target 1 +1/m
               Gives a base mod-3 -1/m to all actions during the duration.
               int 4, psy 4

fumble         cast 3ap 1m, duration 1r +2/m, range 10 +5/m, target 1 +1/m
               Gives base mod-3 -1/m to all actions during the duration.
               int 4, psy 4

pain           cast 1r 1m, duration 3r +2/m, range 10 +5/m, target 1 +1/2m
               pain = diff(cast) or diff(psy vs psy) +2/m
               Determine pain diff version when buying the spell.
               Inflicts pain mods on the target during the duration.
               int 4, psy 6

stun           cast 1r 1m, range 10 +3/m, target 1 +1/m
               Stun 9 +6/m
               int 5, psy 5

frenzy         cast 3r 2m, range 10 +2/m, power +1/m pay at casting
               duration until fail psy vs psy +power each round mod-1/r active
               duration minimum 1 round.
               primary action to roll for keep duration.
               Target is frenzied and will use all actions to attack nearest
               target, all out, regardless of friend or foe.
               int 8, psy 8

earthling      cast 3r 2m, activation range contact, duration 10r +5/m,
               power +3/m
               Gives the target the ability to walk through solid objects
               at a maximum movement speed of maneuver.
               The target must pass a psy+power roll each round or bounce
               back to where he started, taking 1d4 damage penetrating.
               int 5, psy 6

flying         cast 2r 2m, activation range contact, duration 2r +2/m,
               power +3/m
               Gives the target the ability to fly at a minimum speed of run.
               The target must pass a psy+power roll each round or fall down.
               int 6 psy 6

levitate       cast 2r 1m, target 1 +1/m, duration 5r +5/m,
               range group base contact, speed 2 +1/m
               int 4, psy 5

levitate       cast 2r 2m, range 5sq +5/m, duration 5r +5/m, target 1 +1/m
               Gives the targets the ability to levitate slowly at the maximum
               movement of walk. Targets must stay within the range of the
               caster.
               Unwilling targets must be forced by psy vs psy.
               int 6, psy 6

boiler plate   cast 1r 1m, range vision, duration 1r +1/m, target 1 +1/3m
               If the target has a metal armour he takes damage per round equal
               to the abs of the armour, penetrating.
               "Let's boil them in their expensive metal plate..."
               int 9, psy 6

teleport       cast 2r 1m, target 1 +1/2m, range contact.
               Teleport willing targets to any square in line of sight of the
               caster, or to any site the caster has memorised for
               teleportation and/or marked with a teleport marker.
               All transported must roll per, and takes fail diff as mod
               for the rest of the turn.
               Contact with other targets is only necessary for activation,
               not for entire casting.
               A caster can keep int teleportation sites in memory.
               Memorising a site takes 10-per rounds (minimum 1r) and then
               placing a teleport marker on the site.
               A fail-3 will deviate from destination 1d8 direction and
               distance fail^2 sq.
               A fail-6 will teleport the target somewhere else. Way else...
               Materialisation conflicts are resolved in some interesting way.
               A teleport marker costs ~1g and weighs 0.5enc.
               It is also possible to teleport unwilling targets, but it
               requires a psy vs 3psy +1/3m roll and takes 2r instead of 1.
               int 5, psy 5

sentinel       cast 3r 3m, duration 100r +30/m, charge 10 +3/m, range 3
               Creates a forbidding ghost warrior that will attack all non
               friendly targets that approach with a magic ghost spear.
               The sentinel warrior will never move from the square it was
               created. It will make one attack per round at skill 8, and up
               to two parries at skill 8, ignoring ams and initiative.
               The ghost warrior cannot be damaged or parried with non-magical
               weapons.
               It is created with a certain charge. Each attack takes one
               charge, and each hp damage it takes from magical means and
               weapons also costs one charge. It will dissolve when the
               charge is depleted.
               spear 8 dam 4, abs inf, reach 1, initiative 12
               vision 10 360deg dark (3x)
               is light source range 3, soft blue white
               int 8, psy 8, scf 0.66

detect magic   cast 1r 1m, radius 5r +3/m, power per/3 +1/m
               The caster will be aware of all unhidden spells within the
               radius. Hidden spells that are hidden with lower power than
               the detect spell are also detected. Hidden spells that are
               hidden with higher power are not detected. Spells which are
               hidden at equal power are detected 50\% of the time.
               int 5 psy 5

big boss' eye  cast 2r 1m, duration 10r +5/m, move 5sq/r +3/m,
               vision 20 +10/m 360deg +dusk=1m +night=2m
               Creates an almost invisible eye per-6 to notice, which floats
               in the air. The eye cannot pass through solid (1+sq) walls,
               but can pass through doors, thin walls, etc.
               The caster will see everything the eye sees as if he was there.
               Great for getting sight to other spell targets...
               int 9, psy 6

identify       cast 10r 1m, duration 10-int (min 3),
               Identifies up to 1mana spell +1/m, not hidden spells
               Hidden spells at power 0 +1/m
               int 9, psy 5, scf 0.5

disrupt        cast 1r 2m, range 5 +5/m,
               disrupt a spell being cast, requires resistance roll.
               disruptor: psy-3 + 1/2m
               target caster: psy + spell mana
               int 8, psy 8, scf 0.66

dispell        cast 2r 3m, range 3 +3/m,
               dispell an existing spell effect, requires resistance roll:
               dispeller: psy-3 + 1/2m
               target spell: psy(caster) + spell mana
               int 8, psy 3, scf 1.0

distribute     cast 1r 1m, range 5 +5/m,
               move one item +1/m from one willing character to another
               int 5, psy 3

teamspeak      cast 2r 1m, range 10 +10/m, duration 10 +10/m,
               targets int +3/m
               Select targets which will be able to hear each other without
               as long as they are within range of each other (not caster)
               even if they whisper.

boost          cast 1r 1m, range psy +5/m, duration 5 +5/m,
               targets 1 +1/2m
               Increase one of the target's base stats
               (str,dex,con,int,psy,per,cha) by 2 +1/m for the duration.
               Casting is mod+1 if touching target.
               Once cast the target can move out of range



summon demon   ! don't buy this yet, not finished with the background work
               cast 3r 3m, power 0 +3/m, duration 5r +5/m
               Summons a demon for the duration of the spell
               roll list dX: (common demon 5 slots, rare 1 slot)
                1 -  5  grey stalker
                6 - 10  green mouth
               11 - 15
               xx - +1  large grey stalker



perhaps a magic trap can come in handy sometimes
------------------------------------------------

embed magic    cast 3r 1m, duration 100r +100r/m, capacity 5 +5/m
               Can embed another magic formula in an inanimate object to be
               triggered by touch, destruction, or other "mechanical"
               disturbance. The capacity limits the power of the embedded spell.
               int 7, psy 5, scf 0.25

trigger rune   cast 5r 1m, duration 100r +100r/m, range 10 +5/m
               Can trigger one or more embedded formula within range.
               Triggers can be complex combinations of mechanical disturbance,
               sound, light, etc, with complex logic, but very limited
               "intelligence".
               It can react to triggering conditions within its range, and can
               target guide the triggered embedded spells.
               int 7, psy 5, scf 0.25

hide casting   cast 1r 1m, duration 5r +5r/m, capacity 5 +3/m, power 0 +1/m.
               Hides a formula being cast so that it cannot be detected by
               detect magic spells, tingling magic ability, etc.
               The power is set against the sensitivity of the detector.
               The capacity must be equal or larger than the total mana of the
               spell that is to be hidden.
               int 8, psy 8, scf 0.25


\end{verbatim} \normalsize
%\pagebreak[1]
%\small \begin{verbatim}
%TODO embedding stuff etc, NOT COMPLETE %------------------------------------------
%
%hide embedding cast 2r 1m, duration 100r +100r/m, capacity 5 +5/m
%               int 8, psy 8, scf 0.25
%
%permanence     cast 50r 5m, power 0 +1/3m, range 5 +1/m
%               The power must be equal to the total mana cost of all embedded
%               spells to make permanent.
%               Makes spells permanent. Used to embed magic in items.
%               It does not make the power permanent, the embedded permanent
%               spells still require mana supply to function more than once.
%               int 10, psy 10, scf 1.0
%
%storage        cast 25r 3m, power 0 +1/1m, range 5 +1/m
%               The spell creates a magical storage for mana energy. It can be
%               charged and discharged by transfer spells, etc.
%               Used to build and power magical devices.
%               int 10, psy 10, scf 1.0
%
%generate       cast 100r 5m, power 0 + 1/3m, range 5 +1/2m
%               The regen spell will function as a mana supply, generating
%               mana point equal to its power each day, just like a character.
%               Used to build and power magical devices
%               int 12, psy 12, scf 1.5
%
%nexus          cast 1day 10m, power 0 +1/10m, range 5 +1/2m
%               The power is the amount of mana that will be supplied per
%               activation discharge. This is the ultimate mana power supply,
%               since it can activate again and again in rapid succession when
%               combined with permanence.
%               It is susceptible to weak mana flow and void fields though,
%               by supplying lower and perhaps insufficient amounts of mana.
%               The mana coming from a nexus spell fleeting and strange. No one
%               has yet managed to figure out a way to store the mana for later
%               use. It has to be used immediately by channelling it into a
%               spell.
%               int 12, psy 15, scf 2.0
%
%
%
%
%\end{verbatim} \normalsize



\

More spells on the way, not balanced 

\

\small \begin{verbatim}
destructive dispell
    cast 1r, 1m, range 20 +10/m, 
    psy+3/m vs psy+mana
    blast with range of target spell mana
    damage is target spell mana
    i.e. target spell disintegrates in a blast with damage and radius equal to mana

destructive feedback
    cast 1r, 1m, range 20
    psy+3/m vs psy
    damage 6 penetrating and draws equal mana from target, 1mana/damage
    i.e. target's mana is damaging
    Cannot draw/damage below target mana 0.

invisibility
    cast 1r, 1m, personal +1 target/2m touch
    duration 3r +2/m

elsewhere
    cast 3ap, 1m, personal
    teleport 2D4-5 sq X/Y (+/-3sq)
    jump teleport anywhere else, away from incoming blow
    (alternatives: 2D6-7 +/-5sq , 2D3-4 +/-2sq)
    If the target square is occupied the caster bounces back to original square
    but an incoming blow has perhaps missed, treat as avoid success+9 roll
    unmodified.

goaway - knockwave
    cast 1r, 1m, self, radius 1sq+1/m
    knockback 1+1/m, affects all within radius
    all affected are knocked back, away from the caster

fiery eye        int 5 psy 5
    cast 3r 1m, touch, duration 20+1d10r +10/m       (duration based on object?)
    creates fire, light source r10sq +5/m
    the caster gets vision 360deg from the fire source.  (wish for tool to limit
    default day vision, +1m dusk, +1m night, +1m dark     rate of turning of arc
    range to target object is unlimited                   eg 90deg, 45deg/r rate
                                                          of rotation)

one hand
    cast 3r 1m, range 10 +5/m, duration 5r +3/m
    One of the target's limbs is paralysed and useless.
    psy+3/m vs psy each round to maintain spell after first round.
    
other's hand
    cast 3r 1m, range 10 +5/m, duration 5r +3/m
    The caster takes control of one of the target's limbs.
    Any actions taken with the limb are paid from the caster's ap, and are
    slow+1 ap from usual cost. Any action requires a roll to succeed limited by
    the minimum of any skill the action requires and the success diff of the 
    psy vs psy control roll for the round:
    Control: psy+3/m vs psy each round to maintain spell after first round.
    Maintenance: int-1, psy-1

mimic
    cast 2r 1m, range 10r, duration 10r +5/m
    range after casting: 30r +20/m
    Make an object look and feel like something else.
    A perception roll penalty modified by casting success will see that there is
    something strange with the object. Mod by the complexity of the object.
    A find roll will see through the illusion.
    Maintenance: any movement or significant complexity requires int-1 psy-1
    Just stationary object is regular psy-1.

-

\end{verbatim} \normalsize


%--------|---------|---------|---------|---------|---------|---------|---------|
%       10        20        30        40        50        60        70        80
%-------------------------------------------------------------------------------
% removed spells
%
% hasten ?





%-------------------------------------------------------------------------------
%M A G I C A L   E Q U I P M E N T
%---------------------------------

\phantomsection\addcontentsline{toc}{section}{equipment}
\section*{Magical equipment}

Wizards and Warriors alike can gain great power by wielding the arcane artefacts imbued with various magical energies. Below are some suggestions of magical items and artefacts. Make up your own to suit your adventure.

\

\small \begin{verbatim}
Quaff! the     Metal tin with a golden glowing liquid. Generally costs ~1gold.
soft drink     Restores 4 hitpoints and 4 stamina in one round.
of heroes      Has a tendency to be addictive.
               Advertisement: "Yngolf Storpec says it even tastes good."
               The fine print: All side-effects are purely co-incidental
               and have never been scientifically or thaumaturgically proven
               to be caused by the use of Quaff! Quaff corp is not responsible
               for any mental or physical disturbances, changes, manifestations,
               or other interesting or harmful events.

Blast Off      Beautiful crystal vial of clear liquid. Costs ~5gold.
               restores 5+1d5 hitpoints and all stamina in 5 rounds.
               If the vial is broken before empty it detonates doing
               1d5 damage to a radius 1 area blast.
               Everyone caught in the blast get 2 stamina restored.

The Seeking    mod+3, dam 5, range 16, str 3 (no str bonus)
Bow            range mods as regular bow, rate of fire as regular bow,
               The arrows will always curve and re-target to the largest 
               creature within range 10 regardless where it is aimed 
               (except user). Outside range 10 it gives a mod+1 to hitting the
               selected target. Largest is decided by base size, or str.
               Fuck-off Fred swears it seems to like friendly fire.

Slight hammer  dam 5, abs 14, str 5, finesse-2
of solitude    After a regular hit the user can choose to spend 1 mana to:
               roll psy+str vs opponent str. Effect = diff/3
               For every effect point the opponent is dealt 1 damage
               penetrating and knocked back 1 square.
               Charisma -2 while owned (grumpy).

Mostly         dam 7(12), abs 15, str 7, finesse-2
Magical        On each hit, roll for 3 on a d10, if successful the MMM activates
Morningstar    and forcibly draws 1 mana from the user,
(heavy)        same rules for unconsciousness as "casting when dry".
               Activation adds 5 extra to max damage of the weapon (12) for the
               strike that just hit.
               All damage is treated as both physical and magical damage for
               purposes of damaging magical and non magical living creatures.

Terribly       A magical torch with a perpetual soft bioluminescence (range 8).
Tasty          It grows magical glowing mushrooms from the top.
Torch          They are very tasty. Any monsters nearby (distance 3) must 
               succeed a psy roll or get the torch and eat the top mushrooms.
               They grow back in a day, but the torch is dark without them.
               Eating the mushrooms will cause "frenzy" (see spell)
               for 5+1d10 rounds.
               They cost a lot, difficult to find, and illegal in most areas.

The DMT of     The last person who found out what DMT stood for died in the
Dread and      process. It is a box of tools, and it gives off a soft "ping"
Utility        at regular intervals. A hammer is missing.
               It gives a +6 to mechanical, locks, traps rolls, 
               and +3 to McGyverism, find, dungeoneering rolls

The Shiny      The helmet really shines. Provides a 20 distance bright light.
Helmet         It will draw one mana at the start of every 30 rounds.
               It's a scary 5 object against monsters that dwell in darkness.

The One Ring   There is only this one!

Banors Gate    A large orb set in a stretcher-like carrying contraption.
               It requires a total of 15 str to carry around.
               Up to four people can help carry the device.
               The Orb is magically linked to a secondary portal plate.
               Touching the orb activates a teleport to the plate.
               Stepping on the plate activates a teleport to the orb.
               The plate cannot be moved once linked with the orb. That would
               break the link, and the orb must be placed on the plate to
               create a new binding link.
               The orb is charged with a certain numbers of teleport activations
               when it is bound to the plate. Re-charging requires re-binding.
               Teleporting from the orb requires contact with it for five
               rounds. Teleporting from the plate to the orb takes one round.

Metaphysical   A harness with a faintly glowing red tendril sticking out of its
cord of        back, extending about a meter, dissolving into nothing.
Wunjee         Activating the harness takes one round and the wearer is flung
               through meta-space to a anchoring point, taking 1d4 damage
               penetrating. The harness is then physically bound to the
               anchoring point by a thick solid tether 3sq long,
               with abs15 and hp10.
               Climbing out of the harness takes 3r.
               Most are single use, but can be recharged at lower cost than
               a new item.

Metaphysical   A device with chains, hooks, locks, spikes and other interesting
anchor of      protrusions. It must be fastened to something very sturdy.
Wunjee         A red glowing tendril extends for about a meter, dissolving into
               nothing. Fastening the anchor takes 20r, unclasping takes 10r.
               It had better be really securely fastened.
               Most are single use, but can be recharged at lower cost than
               a new item.

Eileens Eye    A glass eye. It looks very realistic.
               The owner of the eye always sees what the eye sees, regardless
               of how far they are apart.
               vision: range 20 arc 180deg.

Wailing Doll   A small child's toy, a rag doll. It is very afraid of monsters.
               When it sees a monster it comes alive and starts screaming very
               loudly. It stops when it doesn't see any more monsters.
               1hp, movement 0.
               vision 5 arc 180deg.

Whaling Doll   A giant child's toy, a wooden construct giant. It requires str20
               to move the doll. Up to four people can help carry it.
               When it sees a monster of str>=10 it comes alive, stands,
               and attacks with a large harpoon spear.
               It makes one attack each 3 rounds (Throw, retract, ready).
               str 30, hp50, abs3, movement 0,
               will turn to face enemy when active
               harpoon 7 dam15, range15
               If hit, the target is bound by the harpoon wire to the doll,
               which will reel it in and try to tear the target apart limb
               from limb (str vs str).
               vision 20 arc 225deg.

Time Fly       A small crystal with a strange looking bug encased.
               Crush the crystal and the bug will turn back time for the user
               by one action. The user may then choose to do something
               else than he tried last time. Perhaps run away?
               The user acts first for the redo action, ignoring initiative.
               Crushing the bug is a 0ap interrupt action ignoring initiative.
               This assumes the bug is in a quickdraw slot, otherwise it must
               already be held or it is useless.
               Time Flies are very exotic and expensive items.
               Ad: A monster charges and attacks. You valiantly defend,
               but get slaughtered! No problem, just Bash the Bug!
               And you are right back at the monster just charging you.
               This time you run away, seeing that you can ignore initiative!
               The (movement +) attack-defend exchange or (movement +)
               attack(miss) - counter attack is considered to take
               place in one action time as long as the bug-crusher has taken
               no further actions.
               weighs 0.1 enc

Stunning       A beautiful white javelin with dark black runes along the shaft.
Sticker        When striking a target and inflicting at least 1hp damage after
               armour the target must roll 10 vs (highest psy or con),
               or be stunned for diff/3 (round up) rounds.
               Even if resisted it gives a stun 6 effect from the shock.
               weapon stats as normal javelin, mod+1 for ranged and melee.

Sentinel Base  A large disc, usually made of stone and metal. When dropped and
               the top seal is broken it will summon a sentinel ghost warrior
               (see the spell). Charge, skill, weapons, and appearance can
               vary between different versions and manufacturers.
               Quite expensive, but very useful sometimes.
               Weighs around around 3.0 - 8.0 enc

Ward Rune      A large tattoo, usually on the back. When the character is
Tattoo         attacked the tattoo flares up and protects him.
               Usually it absorbs 1 damage and draws 1 mana. Various quality
               and power absorbs and draws differently. Usually the wearer
               can choose not to activate the tattoo if he wants to save mana.

Glowrb         A crystal orb which when touched starts to glow with a bright
               white light. It stays alight for 10-1000r or until touched again.
               Single use, charge, or permanent, depending on price and quality.
               enc 3.0 - 5.0

PortaPortal    Two large metal discs, comes as matched pairs. When both are
               placed on the ground and both seals are broken it creates a
               short lived portal between the disks. The portal usually lasts
               for about 10r. A disk must not be moved once the seal has
               been broken and that end of the portal activated.
               Often used to bring in a party to areas which is otherwise
               difficult to get in to, such as a well guarded keep, a remote
               mountain cave, or the inner treasure chamber of the dragon's
               lair. Just have your stealthling sneak in one of the discs of the
               portaportal, then activate. Meanwhile the rest of the party are
               safe and sound back in the tavern with the other end already
               active. Can also be very good for escaping a tight spot. Just
               make sure the monsters don't follow you through.
               Usually activates in 1-3r depending on quality.
               Single use only.
               enc 3.0 - 10.0 x2, usually ~5.0

Amulet of      When the character is taking hp damage the amulet activates and
random         absorbs 1d3 damage, draining one mana in the process.
protection     Some versions protect more, or drain more, or less, mana.
               enc 0.1-0.5

Proportional   The proportional plate mail absorbs 1/3rd of the incoming damage
Plate Mail     (round to nearest). Thus for heavy opponents it can
               be a real life saver, while for smaller critters it is nearly
               useless.

TurBoots       Magical stuffed fish slippers. Spend a mana and increase your
               movement: m+1 w+2 r+4 d+8. Must roll dex at end of movement 
               or fall. Take mod+3 to the roll if going off balance. Or just do 
               a faceplant as a controlled fall.
               Or get the maneuver "TurBoot Shuffle" (5xp) to skip the dex roll.

Feedback Football  angry glowing metal ball contraption
               glow 2sq light, destructuve feedback triggered range 6sq
               6 dam drain equal mana
               triggers on touch (pickup, but not kick)
               triggers on any casting in range, 8 charges
               -- TODO -- alt text --
               activates when thrown, deactivates when packed away, 8 charges
               attack on touch or detecting casting within 6sq: destructive feedback
               draws 1d6 mana for max 6 dam, penetrating : resistance 15 vs psy

Long Sword of  dam 8, abs 14, str 6 (max +2 str bonus)
Irdereng       on successful hit the attacker can choose to roll psy,
               if successful the attack does diff (psy roll) extra damage,
               penetrating, and draws one mana point from the user.
               finesse/poke/swing as usual

Angus' Shield  abs 20, parry+3, str 4
               when declaring parry, the user can choose to spend 1 mana
               and get another parry+3 mod. Only one mana per parry action.

The Very Evil  dam 2 pen 1 abs 5 (no str bonus), parry-3, todefend-3, finesse 9
sacrificial    Starting a ritual sacrifice takes 10r of mumbling mumbo jumbo.
dagger         Drains-converts 1hp to 1m per round (after 10r start).
               Cannot drain from dead/undead targets.

crystal ball   holds 20 mana, transfer 1m/r in/out
               Transfer is full round action, roll psy.
               Must stand still.

energy rod     holds 10 mana, transfer 3m/r out, transfer 1m/3r in
               Transfer is full round action:
               Roll psy, if fail: diff mana is lost in the transfer.

psycharmour    abs 2 +2/m must declare and roll for skill before damage is dealt
               (i.e. after hit, but before to damage rolls) -- not all versions.
               Some versions can declare and roll after damage is rolled.
               Some versions must declare and roll before attack is rolled.
               Defensive mana comes from armour store, not user.
               Defensive mana spending is not an action, it's a 0ap interrupt.
               Holds 15 mana, transfer 1m/2r in, full round action, roll psy.
               psycharmour skill is scf 0.33

powerstaff     dam 3/4 (1h/2h), abs 10, parry 0/+1, str 3 (max +1 str bonus)
               Holds 10m. Can only release 1 point per spell casting, and the
               released mana must be used to reduce the spell mana cost to the
               caster by 1 point. Release is a 0ap free action, included in the
               spell casting, no roll required.
               Transfer to staff 1m/r, full round action, roll psy.
               Cannot transfer power from staff to bearer.
\end{verbatim} \normalsize

\

There are of course also more normal boring magical items, weapons, and artefacts.

\

\small \begin{verbatim}
weapon +x     A weapon that is easier to use or does more damage. It can give
              the user a bonus to his weapon skill, or to his damage roll, or
              to both. Or have exceptional abs resiliance, or ...
              Weapons like this does not have to be magical. Master or Legendary
              craftsmen can make +1 or even +2 weapons.

healing potion
stamina potion
mana potion
[ability] potion
[skill boost] potion
[stat boost] potion
\end{verbatim} \normalsize

\

\clearpage
\TODO custom magical stuff is not finished yet... 

\

\textbf{Just throw out and rewrite!}

\

\emph{Custom magical items: the following are just idea sketches and should not be taken seriously. I need to work on this quite a lot before it is useful.}

Purchasing custom magical potions: A quick approximate cost for general low power magical potions: \\
cost in gold ~ 1 + (0.025 * (mana+2) * scf * psy * int) \\
The extra mana is to approximate the cost of trigger rune and embed magic spells.


\subsection*{Manufacturing custom magical potions}
%-------------------------------------------------
The wizard must know both "embed magic" and "trigger rune" to bind the spell into the potion. Then it requires a suitable potion base liquid and vessel. Those are not so cheap. The vessel and liquid is lost if either the spell, the binding or the trigger rune fails when casting.

The cost of a suitable liquid and vessel can be calculated as: \\
cost in gold ~ 0.25 + (0.01 * mana * scf * psy * int) \\
It is possible to use inferior liquids and vessels, but the potion will lose potency or stop working all together pretty quickly and unreliably.
Don't forget to include the mana cost of the trigger rune and embed magic spells when calculating the vessel cost.

Magical potion: \\
Get the right vessel and liquid, or other containment item.
Then cast the embedding spell and the actual spell. Since a potion is to be quaffed it requires no complex trigger rune, and the very simple triggers in the embed magic spell are enough.
E.g: a potion of fire storm can be created like this:
Fire storm spell is scf 0.33, int 6, psy 8, and the customer wants radius 2 and damage 6, so the spells requires 3 mana by itself. Then an "embed magic" spell is required to bind the fire storm into the potion. When binding into a "suitable material" the duration of the embedding is of no consequence, and thus only the capacity is of interest. The embed spell has a base capacity of 5 and that is enough for the fire storm spell.
The cost of the liquid and vessel is thus approximately:\\
base cost: 0.25 gold \\
fire storm requirements + (0.01 * 3 * 0.33 * 6 * 8) gold \\
embed magic requirements + (0.01 * 1 * 0.25 * 8 * 5 ) gold \\
which in total = 0.25 + 0.48 + 0.1 = 0.83 gold. \\
This is of course an approximation of the cost of the liquid and the flask. Actual prices will vary, and the suitable materials might be difficult to acquire.

Purchasing ready made magical potions is more expensive, with cost approximately \\ 1 + (0.25 * (mana+2) * scf * psy) gold.


\subsection*{Manufacturing custom magical equipment}
%---------------------------------------------------
Manufacturing magical items is more expensive than potions. There is a lot more magic involved and the materials to be enchanted need to be able to hold the magical energies and forces involved.

Single use or limited use magical items can be manufactured similar to the potions, except they usually need a proper trigger spell to focus the spells at the right target.
The approximate cost of materials can be estimated as: \\
1.0 + (0.25 * mana * scf * psy) gold \\
for simple single or multi use items such as wands, crystals, etc.

The interesting stuff, the permanently magical items require



Manufacturing custom magical equipment is similar, though often more expensive, but then the caster also needs to know the "permanence" spell to make the magic last, together with the "nexus", "transfer", or "drain" spells for powering the item. And if any of the spells fail the item will be warped, destroyed and useless.

The required magic to create a single use magical item is just like the magical potion, except it also needs a trigger rune, embedded with the spell.

To create a many use magical item is very similar as a one use, just cast the spell+trigger+embed multiple times, until the items magical capacity is full.

To create a permanent magical item things become more tricky. The "permanence" spell must be cast to encompass the payload spell, the trigger, mana source, and embedding. And of course they also require some form of mana source.

Mana sources come in different forms. The simplest is that the user of the item knows the "magic" skill to sufficient level to power the payload spell with his own mana. This is not so handy though, since most customers don't have the magic skill, and thus cannot empower magical spells.
The next step up is to embed a "mana transfer" spell together with the payload spell. This way the item can draw mana from a willing user.
The next step up is to embed a "storage" spell with the payload with sufficient mana to use the item once or a few times, before the storage needs to be recharged.
The next step up is to combine the storage with a "generate" spell which slowly replenishes the storage spell over time. This way the item will recharge by itself, and can still be manually recharged by a mana transfer spell if needed.
The coolest and ultimate mana supply is the "nexus" spell, which always can supply mana directly to the payload spell with high frequency ad infinitum.

Here follows an example throwing knife of slow motion \\
Payload: slow: mana 3 (+2m duration), scf 0.33, int 4, psy 3 \\
Trigger: trigger rune: mana 1 (contact with target), scf 0.25, int 9, psy 5 \\
Supply: storage: 12 mana (+9m capacity), scf 1.0, int 10, psy 10 \\
Supply: generate: 14 mana (+3m recharge), scf 1.0, int 12, psy 12 \\
Embed: embed magic: 7 mana (+6m capacity), scf 0.25, int 8, psy 5 \\
Permanence: permanence: 116 mana (+111m capacity), scf 1.0, int 10, psy 10 \\
Which creates a throwing knife of slow motion, which can be used three times on its own power, and recharges one use per day.

Now, the approximate cost of suitable material for the throwing knife would be something like this: \\
payload: 0.01 * 3 * 0.33 * 4 * 3 = 0.12 \\
trigger: 0.01 * 1 * 0.25 * 9 * 5 = 0.11 \\
supply: 0.01 * 12 * 1.00 * 10 * 10 = 12 \\
supply: 0.01 * 14 * 1.50 * 12 * 12 = 30 \\
embed: 0.01 * 7 * 0.25 * 8 * 5 = 0.70 \\
The permanence requirements does not fall on the target object, but rather the equipment of the wizard creating the item. He must have such an object available to build up the permanence spell in before sealing the spells of the throwing knife. \\
permanence: 0.01 * 116 * 1.0 * 10 * 10 = 116 \\
Thus the materials requirements for the knife will probably cost somewhere around 40-45 gold, at break even for the manufacturer! Expect the wizard to require additional funds for the work and whatnot, and of course the weapon smith will require quite a bit to custom make the knife itself, before it is handed over to the wizard to be enchanted.

The same knife, but drawing mana from the user is much cheaper to make: \\
Payload: slow: mana 3 (+2m duration), scf 0.33, int 4, psy 3 \\
Trigger: trigger rune: mana 1 (contact with target), scf 0.25, int 9, psy 5 \\
Supply: mana transfer: 1 mana (get 3 mana), scf 0.33, int 3, psy 3 \\
Embed: embed magic: 1 mana (5 capacity), scf 0.25, int 8, psy 5 \\
And the permanence requirement \\
Permanence: permanence: 35 mana (+30m capacity), scf 1.0, int 10, psy 10 \\
Which gives a total estimate of the knife material cost at: \\
payload: 0.01 * 3 * 0.33 * 4 * 3 = 0.12 \\
trigger: 0.01 * 1 * 0.25 * 9 * 5 = 0.11 \\
supply: 0.01 * 1 * 0.33 * 3 * 3 = 0.03 \\
embed: 0.01 * 1 * 0.25 * 8 * 5 = 0.10 \\
permanence: 0.01 * 35 * 1.0 * 10 * 10 = 35 \\



Magical capacity of weapons and items:\\
capacity factor of material * abs * encumbrance / 10.

steel:       5
star iron:  20
mage steel: 10
mithril:    50
copper:     15, brass 10
silver:     20, alloy 15
gold:       30, alloy 20
crystal:    50
gems:       1/gold value



\

\TODO: snipped moved from prices.tex :

Permanently magical items are always custom made, and very exotic equipment.
First of all the item to be enchanted needs to have enough magical capacity to withstand the forces of the magic to be embedded within it. This means that normal steel will almost never suffice, and more exotic and expensive materials will have to be used in the making of the item itself.

The magical capacity of an item can be calculated to approximately: \\
capacity = cf * abs \\
where cf is capacity factor of material and the skill of the maker, the abs is the absorption or hitpoints of the item.
E.g: a steel sword of regular making has a magical capacity of 10 = (1 * 10).

Here follows the magical capacity of some materials:

\

\small \begin{verbatim}
material          capacity      abs  enc  dam   price

wood              1.0    1      1.0  1.0  1.0    1.0   (for common use)
gnarlwood         5.0    5      1.0  1.0  1.0   10.0   (instead of wood)
steel             1.0    1      1.0  1.0  1.0    1.0   (for common use)
mage steel        5.0    5      1.0  1.0  1.0   10.0   (instead of steel)

copper alloy      9.0   15      0.5  1.2  0.8   30.0
silver alloy            25      0.4  1.5  0.7   50.0
gold alloy              40      0.3  2.0  0.6  100.0
strong crystal          50      0.5  1.0  1.2
mithril                         2.0  0.7  1.2
star iron

copper                     ~3/enc
silver                     ~5/enc
gold                      ~10/enc
jewels            1*cost ~100/enc

\end{verbatim} \normalsize

The capacity requirements of a permanently magical item are:
payload spell: cap = mana * psy (E.g: black bolt 3m = 21 capacity)
trigger spell: cap = 5 per payload spell for normal complexity triggers
no mana supply: cap = 0 (empower by user "magic" spell)
mana transfer: cap = 1 per mana (payload requirement)
mana storage: cap = 30 + 10 per mana
mana generation: cap = 50 + 30 per mana
mana nexus: cap = 100 + 100 per mana
embedding: cap = 1 per mana (total mana)

Permanence doesn't require cap of the target object, but from the ritual equipment of the wizard.






%-------------------------------------------------------------------------------
%F A M I L I A R S ,   M A G I C A L   C O M P A N I O N S
%---------------------------------------------------------

\phantomsection\addcontentsline{toc}{section}{familiars}
\section*{Magical familiars and companions}

Familiars and other magical companions are more intelligent than trained animals and often share a mind link with the owner.





%--------|---------|---------|---------|---------|---------|---------|---------|
%       10        20        30        40        50        60        70        80
